---
title: "Lab 6"
author: "Chase Tarr"
date: "11/1/2021"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a Rank Based Test
 - We will make our own samples using a pseudo random generator]
 - using set.seed (starting point value in a vector that gets created of random values)
    - set.seed tells R to create a sample of integer values
 
```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # values between 0 and 20, with 15 total values

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE) # replace allows the same number to be pulled multiple times (replaced back into pool of values)
```
 
```{r}
hist(gp_1)

hist(gp_2)
```

Try a t-test:
```{r}
t.test(gp_1, gp_2)
```

- This p-value means that if these samples were drawn from populations with the same mean, the probability of taking two random samples with means *at least as different* as the sample means as we found by random chance (taking into account spread, n) is 19.8%. 

- It is likely that the null hypothesis is true so we retain or fail to reject the null hypothesis. >> there is no significant difference in means from group 1 to group 2. (never say "accept")


Now lets compare this outcome to a rank-based test

## Mann Whitney U unpaired rank-based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

- This p-value means that if these samples were drawn from populations with the same ranks, the probability of taking two random samples with ranks *at least as different* as the sample ranks in our samples is 28%. 


